---
title: "Get Final Data Object"
author: "Keith Mitchell"
date: "3/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(reshape2)
```

### First we will start with the GDP and do a melt to get one column representing quarter and one quarter representing
```{r}
gdp_data = read.csv('../Data/Matched Data/GDP Data matched.csv')
gdp_data = melt(gdp_data, id.value="Period", measure.vars=colnames(gdp_data)[2:length(colnames(gdp_data))])
colnames(gdp_data) = c('Country', 'Quarter', 'GDP')
head(gdp_data)
unique(gdp_data$Quarter)
```

### Next we will merge with the Covid by quarter object, here we will keep this date format but generate the format from previous dataframe as well in order to merge on
- In the future this should be generalized to automatically detect and strip at least the year for automation as years increase or generalization are of interst. `strplit` and `paste` just too lazy rn :)
```{r}
data = read.csv('../Data/Matched Data/CovidbyQuarter.csv')
head(data)
unique(data$Date_reported)
```
### Lets get the date values from the covid counts that will associate with quarter before it?
- TODO: check with the group if this should be the date the quarter starts.
```{r}

quarter_value_list = c()
for (i in seq(1, length(rownames(data)))){
  
  if (data$Date_reported[i] == '1/3/2020'){
    quarter_value_list = c(quarter_value_list, 'Q4.2019')
  }
  else if (data$Date_reported[i] == '4/3/2020'){
    quarter_value_list = c(quarter_value_list, 'Q1.2020')
  }  
  else if (data$Date_reported[i] == '7/3/2020'){
    quarter_value_list = c(quarter_value_list, 'Q2.2020')
  }
  else if (data$Date_reported[i] == '10/3/2020'){
    quarter_value_list = c(quarter_value_list, 'Q3.2020')
  }
}

data$Quarter = quarter_value_list
head(data)
```


### Lets merge the last two dataframes (GDP per quarter and Covid cases per quarter)
```{r}
gdpcovid_quarter = merge(x = data, y = gdp_data, by = c("Quarter", "Country"), all.x = TRUE)
head(gdpcovid_quarter)
```

### This data is monthly over the two year so lets just take the values representing the quarters so we merge this into the data frame. Let look at what values we have to work with here
```{r}
unemp_data = read.csv('../Data/Matched Data/Unemployment.csv')
head(unemp_data)
unique(unemp_data$TIME)
```
### Lets convert the months of interest
- TODO: check with the group if this should be the date the quarter starts.
```{r}
quarter_value_list = c()
index_list = c()
for (i in seq(1, length(rownames(unemp_data)))){
  if (unemp_data$TIME[i] == '2020-01'){
    quarter_value_list = c(quarter_value_list, 'Q4.2019')
    index_list = c(index_list, i)

  }
  else if (unemp_data$TIME[i] == '2020-04'){
    quarter_value_list = c(quarter_value_list, 'Q1.2020')
    index_list = c(index_list, i)

  }
  else if (unemp_data$TIME[i] == '2020-07'){
    quarter_value_list = c(quarter_value_list, 'Q2.2020')
    index_list = c(index_list, i)

  }
  else if (unemp_data$TIME[i] == '2020-10'){
    quarter_value_list = c(quarter_value_list, 'Q3.2020')
    index_list = c(index_list, i)
  }
}

unemp_data = unemp_data[index_list,]
unemp_data$Quarter = quarter_value_list
head(unemp_data)
```

###

```{r}
dim(gdpcovid_quarter)
dim(unemp_data)

unempgdpcovid_quarter = merge(x = unemp_data, y = gdpcovid_quarter, by = c("Quarter", "Country"), all=TRUE)#, all.x = TRUE)
unempgdpcovid_quarter_cleaned = merge(x = unemp_data, y = gdpcovid_quarter, by = c("Quarter", "Country"))
dim(unempgdpcovid_quarter)
dim(unempgdpcovid_quarter_cleaned)

unempgdpcovid_quarter
unempgdpcovid_quarter_cleaned
```

```{r}
write.csv(unempgdpcovid_quarter, '../Data/final_data_quarter_outer_join.csv')
write.csv(unempgdpcovid_quarter_cleaned, '../Data/final_data_quarter_inner_join.csv')'
```


```{r}
sessionInfo()
```