---
title: "ModelExplore"
author: "Keith Mitchell"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r warning=FALSE}
library(ggplot2)
library(MASS)
library(tidyverse)
library(kableExtra)
```



### First lets look at the data of interest:
```{r}
inner_data = read.csv("../Data/final_data_quarter_inner_join3.csv")
inner_data$X.unem..TOTAL.[is.na(inner_data$X.unem..TOTAL.)] <- 0
inner_data$X.GDP..TOTAL.[is.na(inner_data$X.GDP..TOTAL.)] <- 0

#inner_data = inner_data %>% filter(!is.na(X.unem..TOTAL.)) %>% filter(!is.na(X.GDP..TOTAL.))

#inner_data #%>% kable %>% kable_styling()
#dim(inner_data)
```


# Full unemp model with Country
```{r}
model11 = lm(X.unem..TOTAL. ~ New_cases_pc + Country, data=inner_data)
modsum = summary(model11)
as.data.frame(modsum[4]$coefficients) #%>% kable %>% kable_styling()
```

```{r}
par(mfrow=c(2,2))
plot(model11)
```


# Full unemp model without country
```{r}
model11 = lm(X.unem..TOTAL. ~ New_cases_pc, data=inner_data)

modsum = summary(model11)
as.data.frame(modsum[4]$coefficients) #%>% kable %>% kable_styling()
```


```{r}
par(mfrow=c(2,2))
plot(model11)
```


# Full GDP model with country
```{r}
model11 = lm(X.GDP..TOTAL. ~ New_cases_pc + Country, data=inner_data)

modsum = summary(model11)
as.data.frame(modsum[4]$coefficients) %>% kable %>% kable_styling()
```

```{r}
par(mfrow=c(2,2))
plot(model11)
```




# Full GDP model without country
```{r}
model11 = lm(X.GDP..TOTAL. ~ New_cases_pc, data=inner_data)

modsum = summary(model11)
as.data.frame(modsum[4]$coefficients) %>% kable %>% kable_styling()
```

```{r}
par(mfrow=c(2,2))
plot(model11)
```


# Box cox procedure was considering but would require a shift in the response variable (GDP and umep total) which would potentially hurt the analysis thus far. Lets try to remove zero values and see if that return nicer QQ plots. 

